---
title: "ibutton data"
author: "Alyssa-Lois Gehman"
date: "2/16/2018"
output: html_document
---

packages and requirements

```{r}
library(plyr)
library(dplyr)
library(lattice)
library(readr)
library(MuMIn)
library(tidyr)

#set working directory so that all files are saved to one location
setwd("~/Dropbox/UBC projects/Hakai/ibutton Hakai temp data 2017")

```

*How to prepare ibutton files for below processing*

1. Graph & look for outliers
2. Filter out unreasonably extreme values
3. Remove poor data days/times as identified by field logs, etc.
4. Match date-time to tide based on logger tidal height--usually as 0 for immersed & 1 for emersed. Using NOAA or Environment Canada observed if available and predictions if not. "Trim" questionable times when water level transitioning if it's important to separate air and water temps.
    canadian tide site:
        http://tides.gc.ca/eng/data/predictions/2017
        
    code for new column labeled 'in.water'    
        0= immersed - in water
        1= emersed - out of water
        2= transition
        
5. Remove the ibutton metadata from the top of the file, resave with a new name in a single folder
    change column label for date to 'date.time'
    For the below code to run you need the column labels to read:
            date.time = col_character(),
            Unit = col_character(),
            Value = col_double(),
            `in.water` = col_integer()
6. Place all processed files into a single folder

*Output of below code*
Three files with these variables:
      avgair= the average air temp (not suggested to be used for analysis - just as a reference)
      maxair=the absolute max air temp (could be used for analysis but generally this could be influenced by outliers)
      sdair=the standard deviation for variation in air temperature experienced
      air90th= the 90th quantile of air temperature experienced
      air99th=the 99th quantile of air temperature experienced (a better max temperature estimate)
      air10th=the 10th quantile of air temperature experienced
      
      avgwater= the average water temp (not suggested to be used for analysis - just as a reference)
      maxwater=the absolute max water temp (could be used for analysis but generally this could be influenced by outliers)
      sdwater=the standard deviation for variation in water temperature experienced
      water90th= the 90th quantile of water temperature experienced
      water99th=the 99th quantile of water temperature experienced (a better max temperature estimate)
      water10th=the 10th quantile of water temperature experienced

```{r ibutton dataset creation}
#remove all lists
rm(list=ls(all=TRUE)) 

#reads all the files in a single folder that have a .csv extension (change so that it goes to your files)
fileNames <- Sys.glob("~/Dropbox/UBC projects/Hakai/ibutton Hakai temp data 2017/processed ibutton data/ibuttoncsvfiles/*.csv")

#creates a list of the file names so that you can label them appropriately below (change so that it goes to your files)
filename<-list.files(path="~/Dropbox/UBC projects/Hakai/ibutton Hakai temp data 2017/processed ibutton data/ibuttoncsvfiles", pattern="csv")

##create empty vectors
f.air=NULL
f.water=NULL
airtemp=NULL
watertemp=NULL

# we'll be leaving the original files unaltered but using their data to create new files:
for(i in 1:37){
  sample <- read.csv(fileNames[i],
    header = TRUE)
  
  #seperate file by water and air
  f.air<-sample %>% filter(in.water=="1")
  f.water<-sample %>% filter(in.water=="0")
  
  ## creating columns to label sites and replicates based off our file name extension: 'BSP.1 iButton Data.csv', or "site.plot iButton Data.csv"
  
  #a column the length of water or air data that names which file the data came from
  fileair<-rep(filename[i],length(f.air[,1]))
  filewater<-rep(filename[i],length(f.water[,1]))
  
  #pull out station (col1) and plot number (col2)
  sair1<-unlist(lapply(strsplit(as.character(fileair), " "),function(z) z[1]))
  swater1<-unlist(lapply(strsplit(as.character(filewater), " "),function(z) z[1]))
  
  sair2<-strsplit(as.character(sair1), "[.]")
  siteair<-do.call(rbind, sair2)
  
  swater2<-strsplit(as.character(swater1), "[.]")
  sitewater<-do.call(rbind, swater2)
  
  #combine site label with ibutton data
  f2.air<-cbind(siteair, f.air)  
  f2.water<-cbind(sitewater, f.water)
  
  #stack the sites in rows
  airtemp<-rbind(airtemp, f2.air)
  watertemp<-rbind(watertemp, f2.water)
  }

#check out the files to make sure that the labeling is done correctly
View(airtemp)
View(watertemp)

#####sumarize by site

## Grouped by site
sum_air_site<-airtemp %>%
            group_by(airtemp[,1]) %>%
                summarise(avgair=mean(Value), maxair=max(Value), sdair=sd(Value), air90th=quantile(Value, 0.90), air99th=quantile(Value, 0.99), air10th=quantile(Value, 0.10))

sum_water_site<-watertemp %>%
                  group_by(watertemp[,1]) %>%
                    summarise(avgwater=mean(Value), maxwater=max(Value), sdwater=sd(Value),water90th=quantile(Value, 0.90), water99th=quantile(Value, 0.99), water10th=quantile(Value, 0.10))

sumbysite<-cbind(sum_air_site,sum_water_site[,2:4])
##combine and write summary file to csv
#write.csv(sumbysite, "temp_airwater_bysite_5Mar18.csv")

###### summarize by site and plot
#grouped by site and by plot
sum_air_plot<-airtemp %>%
              group_by(airtemp[,1], airtemp[,2]) %>%
                    summarise(avgair=mean(Value), maxair=max(Value), sdair=sd(Value), air90th=quantile(Value, 0.90), air99th=quantile(Value, 0.99), air10th=quantile(Value, 0.10))
#write summary to csv
#write.csv(sum_air_plot, "temp_air_byplot_5Mar18.csv")


sum_water_plot<-watertemp %>%
              group_by(watertemp[,1], watertemp[,2]) %>%
                  summarise(avgwater=mean(Value), maxwater=max(Value), sdwater=sd(Value), water90th=quantile(Value, 0.90), water99th=quantile(Value, 0.99), water10th=quantile(Value, 0.10))
#write summary to csv
#write.csv(sum_water_plot, "temp_water_byplot_5Mar18.csv")


## plots of summary to evaluate for outliers 

xyplot(air99th~`airtemp[, 1]`, group=`airtemp[, 1]`, data=sum_air_plot, pch=16, jitter.x=TRUE)
xyplot(air90th~`airtemp[, 1]`, data=sum_air_plot, group=`airtemp[, 1]`, pch=16, jitter.x=TRUE)
xyplot(air10th~`airtemp[, 1]`, group=`airtemp[, 1]`, data=sum_air_plot, pch=16, jitter.x=TRUE)
xyplot(sdair~`airtemp[, 1]`, group=`airtemp[, 1]`, data=sum_air_plot, pch=16, jitter.x=TRUE)


xyplot(water99th~`watertemp[, 1]`, group=`watertemp[, 1]`, data=sum_water_plot, pch=16, jitter.x=TRUE)
xyplot(water90th~`watertemp[, 1]`, group=`watertemp[, 1]`, data=sum_water_plot, pch=16, jitter.x=TRUE)
xyplot(water10th~`watertemp[, 1]`, group=`watertemp[, 1]`, data=sum_water_plot, pch=16, jitter.x=TRUE)
xyplot(sdwater~`watertemp[, 1]`, group=`watertemp[, 1]`, data=sum_water_plot, pch=16, jitter.x=TRUE)


```

